---
title: "everybody.codes.q5"
format: revealjs
editor: visual
---

```{r message=FALSE, warning=FALSE}
library(here)
library(tidyverse)
```

# Everybody Codes 2025

Information about each challenge can be found at everybody.codes

## Quest 5

### Part 1

Creating Splines to determine the quality of swords

1. Check all segments of the spine, starting from the top.
2. If your number is less than the one on the spine segment and the left side of the segment is free - place it on the left.
3. If your number is greater than the one on the spine segment and the right side of the segment is free - place it on the right.
4. If no suitable place is found at any segment, create a new spine segment from your number and place it as the last one.

```{r}
notes_5.1 <- "24:6,8,2,7,4,3,1,8,7,8,3,3,7,4,1,6,1,9,2,9,1,6,7,3,5,4,6,7,8,1"

# extract all values into vector
vals_5.1 <- as.numeric( strsplit( str_extract(notes_5.1, "(?<=:).+"), ",")[[1]] )

```


```{r}
spine_fx <- function(vals_vec) {
  
  S <- c(vals_vec[1])
  L <- c()
  R <- c()
  
  for(i in 2:length(vals_vec)) {
    
    print(paste("L", L))
    print(paste("S", S))
    print(paste("R", R))
    print(vals_vec[i])
    
    if(vals_vec[i] < S[length(S)] & length(L) < length(S) ) {
      L <- append(L, vals_vec[i])
      }
    else if(vals_vec[i] > S[length(S)] & length(R) < length(S) ) {
      R <- append(R, vals_vec[i])
      }
    else {S <- append(S, vals_vec[i])}
    
  }
  
  L <- c(L, rep(NA, length(S) - length(L)))
  R <- c(R, rep(NA, length(S) - length(R)))
  
  results <- data.frame(L, S, R)
  
  return(results)
}

q5.1 <- spine_fx(vals_5.1)

print(q5.1$S)

paste(q5.1$S, collapse = "")
```


```{r}
test <- c(1,2)
test[3]

test <- append(test, 4)

append(test, 5)

test1 <- c(6,7)

cbind(test, test1)

test[length(test)]
```


```{r}
df_vals <- cbind(x = seq(1:length(vals_5.1)),
                 y = vals_5.1,
                 position = rep("i", )

df
```



