---
title: "everybody.codes.q12"
author: "Ben Jefferies"
date: "2025-11-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(here)
library(tidyverse)
```

# Everybody Codes 2025

Information about each challenge can be found at everybody.codes

## Quest 12

### Part 1

How many barrels ignite? You light the top left corner barrel. A barrel can ignite any barrel to the top, bottom, left or right that is equal or less value.

```{r}
# convert data into a numeric matrix representing the barrels array
dat_12.1 <- readLines(con = here("everybody_codes_e2025_q12_p1.txt"))

barrel_rows <- sapply(dat_12.1, FUN = function(x){
  strsplit(x, "")
})  

barrels_m <- do.call(rbind, barrel_rows)

rownames(barrels_m) <- NULL

barrels_m <- apply(barrels_m, MARGIN = 2, FUN = as.numeric)

```


code ideas:

create row and col vectors representing all possible positions in barrel array

create row and col vectors to store values of ignited barrels

while loop (boolean something happened)

check barrel_m[ignite]




```{r}
m_row <- seq(1:nrow(barrels_m))
m_col <- seq(1:ncol(barrels_m))

# create data frame with all coordinates
barrels_coords <- expand.grid(row = m_row, col = m_col)


ignite <- 

something_happened <- TRUE

while(something_happened){
  
  
  
}

```

