---
title: "everybody.codes.q14"
author: "Ben Jefferies"
date: "2025-11-27"
output: html_document
---


```{r message=FALSE}
library(here)
library(tidyverse)
```

# Everybody Codes 2025

Information about each challenge can be found at everybody.codes

## Quest 14

### Part 1




```{r}
# read in file as vector with 21 elements, each of which is 
dat_14.1 <- readLines(con = here("everybody_codes_e2025_q14_p1.txt"))

# split characters
m_14.1 <- matrix( unlist(strsplit(dat_14.1, "")) ,
        nrow = length(dat_14.1), ncol = length(dat_14.1),
        byrow = T)

# replace values with 1 and 0
m_14.1[m_14.1 == "#"] <- 1
m_14.1[m_14.1 == "."] <- 0

# convert matrix to numeric for easy calculating
m_14.1 <- apply(m_14.1, MARGIN = 2, as.numeric)

vector_14.1 <- as.numeric(t(m_14.1))



vector_14.1[]


m_14.1[-1,5]
```


Code Ideas:

Each round sum the matrix.  Add sum to total variable.

To check for next round positioning. Need to check each value in the matrix.
Values to check are diagonal. Meaning you check 4 positions, the position +/- rowlength+1 and rowlength-1.  

For this part, I tried to do the calculation as a vector. However, subsetting vectors is strange since for the starting values you would be checking vector[-8], but in R that -8 returns the whole vector but without the 8th element...


```{r}

active_tiles <- function(tiles_vec, row_len) {
  
  total <- 0
  
  #establish check distances
  topLeft <- -1*(row_len+1)
  topRight <- -1*(row_len-1)
  bottomLeft <- row_len-1
  bottomRight <- row_len+1
  
  # loop that determines value of each element for next round
  for(i in 1:length(tiles_vec)){
    
    if( (i + topLeft | i + topRight | i + bottomRight | i + bottomLeft) < 0 )
    
    
    if( sum( tiles_vec[i + topLeft],
             tiles_vec[i + topLeft],
             tiles_vec[i + topLeft],
             tiles_vec[i + topLeft], na.rm = T) )
    
  }
  
}

```




