---
title: "everybody.codes.q2"
format: html
editor: visual
---

```{r message=FALSE}
library(here)
library(tidyverse)
```

# Everybody Codes 2025

Information about each challenge can be found at everybody.codes

## Quest 2

### Part 1

Complex Number Operations. Addition, Multiplication and Division.

\[X1,Y1\] + \[X2,Y2\] = \[X1 + X2, Y1 + Y2\]

\[X1,Y1\] \* \[X2,Y2\] = \[X1 \* X2 - Y1 \* Y2, X1 \* Y2 + Y1 \* X2\]

\[X1,Y1\] / \[X2,Y2\] = \[X1 / X2, Y1 / Y2\]

**Dividing always rounds down (cut off all fractional parts)**

Begin by setting the result to \[0,0\] . Then, complete three cycles of the following operations: \* Multiply the result by itself. \* Divide the result by \[10,10\] . \* Add A to the result.

```{r}
# read in .txt file with names and codes
dat_2.1 <- read.table(here::here("everybody_codes_e2025_q02_p1.txt"),
                    sep = "")

# extract x and y values from string, put in data frame with 2 cols x and y
A <- data.frame( x = as.numeric( str_extract(dat_2.1$V1, "([0-9]+)(?=,)") ),
                 y = as.numeric( str_extract(dat_2.1$V1, "(?<=,)[0-9]+") ) )
```

```{r}

multiply <- function(complex_1, complex_2) {
  
  x1 <- complex_1$x * complex_2$x - complex_1$y * complex_2$y
  
  y1 <- complex_1$x * complex_2$y + complex_1$y * complex_2$x
  
  R <- data.frame(x = x1, y = y1)
  
  return(R)
  
}

Rx <- data.frame(x = 1, y = 3)

multiply(Rx, Rx)
```

```{r}
divide <- function(complex_1, complex_2) {
  
  x1 <- floor( complex_1$x / complex_2$x )
  
  y1 <- floor( complex_1$y / complex_2$y )
  
   R <- data.frame(x = x1, y = y1)
  
  return(R)
  
}

divide(A, Rx)

```

\[X1,Y1\] + \[X2,Y2\] = \[X1 + X2, Y1 + Y2\]

```{r}
add <- function(complex_1, complex_2) {
  
  x1 <- complex_1$x + complex_2$x
  
  y1 <- complex_1$y + complex_2$y
  
  R <- data.frame(x = x1, y = y1)
  
  return(R)
}

add(A, Rx)
```

Begin by setting the result to \[0,0\] . Then, complete three cycles of the following operations: \* Multiply the result by itself. \* Divide the result by \[10,10\] . \* Add A to the result.

```{r}
R <- data.frame( x = 0, y = 0 )

Atest <- data.frame( x = 25, y = 9)

# loop to run 3 cycles of the mathematical process specified
for(i in 1:3) {
  
  print(paste(R, i, "0"))
  
  R <- multiply(R, R)
  
  print(paste(R, i, "1"))
  
  R <- divide(R, data.frame(x = 10, y = 10))
  
  print(paste(R, i, "2"))
  
  R <- add(R, A)
  
  print(paste(R, i, "3"))
  
}

R
```





### Part 2

Complicated algorithm

A=[-3284,68783]

```{r}
Result <- data.frame(x = 0, y = 0)

# bottom left point in grid of breast plate
A <- data.frame(x = -3284, y = 68783)

# top right point in grid of breast plate (add 1000, 1000)
B <- add(A, data.frame(x=1000,y=1000))

# all grid coordinates
grid_x <- seq(from = A$x, to = B$x, length.out = 101)

grid_y <- seq(from = A$y, to = B$y, length.out = 101)

# data frame with all grid coordinates (all combos of grid_x and grid_y)
grid <- expand.grid(x = grid_x, y = grid_y)
                         
```




```{r}

check_points <- function(points_df, initial_R) {
  
  for(pt in 1:nrow(points_df)){
    
    # re-initialize R for each point to check
    R <- initial_R
    
    for(i in 1:100) {
      
      R <- multiply(R, R)
      R <- divide(R, data.frame(x = 100000, y = 100000))
      R <- add(R, points_df[pt,])
      
    }
    
    engraved_pt <- if{}
    
    
    
  }
  
}

for(i in 1:100) {
  
  
  
}
```











