---
title: "everybody.codes.q2"
format: html
editor: visual
---


```{r message=FALSE}
library(here)
library(tidyverse)
```

# Everybody Codes 2025

Information about each challenge can be found at everybody.codes

## Quest 2

### Part 1

Complex Number Operations. Addition, Multiplication and Division.

[X1,Y1] + [X2,Y2] = [X1 + X2, Y1 + Y2]

[X1,Y1] * [X2,Y2] = [X1 * X2 - Y1 * Y2, X1 * Y2 + Y1 * X2]

[X1,Y1] / [X2,Y2] = [X1 / X2, Y1 / Y2]

**Dividing always rounds down (cut off all fractional parts)**

Begin by setting the result to  [0,0] . Then, complete three cycles of the following operations:
* Multiply the result by itself.
* Divide the result by  [10,10] .
* Add  A  to the result.

```{r}
# read in .txt file with names and codes
dat_2.1 <- read.table(here::here("everybody_codes_e2025_q02_p1.txt"),
                    sep = "")

# extract x and y values from string, put in data frame with 2 cols x and y
A <- data.frame( x = as.numeric( str_extract(dat_2.1$V1, "([0-9]+)(?=,)") ),
                 y = as.numeric( str_extract(dat_2.1$V1, "(?<=,)[0-9]+") ) )
```

```{r}

multiply <- function(complex_1, complex_2) {
  
  x1 <- complex_1$x * complex_2$x - complex_1$y * complex_2$y
  
  y1 <- complex_1$x * complex_2$y + complex_1$y * complex_2$x
  
  R <- data.frame(x = x1, y = y1)
  
  return(R)
  
}

Rx <- data.frame(x = 1, y = 3)

multiply(Rx, Rx)
```



```{r}
divide <- function(complex_1, complex_2) {
  
  x1 <- floor( complex_1$x / complex_2$x )
  
  y1 <- floor( complex_1$y / complex_2$y )
  
   R <- data.frame(x = x1, y = y1)
  
  return(R)
  
}

divide(A, Rx)

```
[X1,Y1] + [X2,Y2] = [X1 + X2, Y1 + Y2]

```{r}
add <- function(complex_1, complex_2) {
  
  x1 <- complex_1$x + complex_2$x
  
  y1 <- complex_1$y + complex_2$y
  
  R <- data.frame(x = x1, y = y1)
  
  return(R)
}

add(A, Rx)
```

Begin by setting the result to  [0,0] . Then, complete three cycles of the following operations:
* Multiply the result by itself.
* Divide the result by  [10,10] .
* Add  A  to the result.

```{r}
R <- data.frame( x = 0, y = 0 )

# loop to run 3 cycles of the mathematical process specified
for(i in 1:3) {
  
  print(R)
  
  R <- multiply(R, R)
  
  R <- divide(R, data.frame(x = 10, y = 10))
  
  R <- add(R, A)
  
}

R
```

