---
title: "everybody.codes.2025"
format: html
editor: visual
---

```{r message=FALSE}
library(here)
library(tidyverse)
```


# Everybody Codes 2025

Information about each challenge can be found at everybody.codes

## Quest 1

### Part 1

```{r}
# read in .txt file with names and codes
dat_1 <- read.table(here::here("everybody_codes_e2025_q01_p1.txt"),
                    sep = "")

names_1 <- as.data.frame(str_split(dat_1[1,], pattern = ","),
                         col.names = "names")

codes_1 <- as.data.frame(str_split(dat_1[2,], pattern = ","),
                         col.names = "codes")

codes_test <- codes_1$codes[1:6]
```

For this first challenge, the 11 codes instruct how to move through the list of names provided. For instance, L6 means move Left (up) 6 places. We start at the first element in list (or vector for R's purposes), and move from there. 

Psuedocode: 
* need a number representing the number of positions in the names vector. Limit is 10, so bounds are 1 and 10.
* for each code, check if begins with R or L, R means add number, L means subtract
* extract integer in second place of code and add or subtract
* check at end of each turn if index is < 1 or > 10, if so, set to 1 or 10

```{r}
alg_1 <- function(names, codes) {
  
  index <- 1L
  
  for(i in 1:length(codes)) {
    
     # move index according to code
    index <- ifelse( substr(codes[i], 
                   start = 1, 
                   stop = 1) == "R", 
           index + as.numeric(substr(codes[i], 2,2)),
            index - as.numeric(substr(codes[i], 2,2)))
    
    if(index < 1) {
      index <- 1
    }
    if(index > length(names)) {
      index <- length(names)
    } 
    else {index <- index}
    
  }
  
  return(names[index])
  
}


alg_1(names_1$names, codes_1$codes)


```




### Part 2

To find your parents names, follow a code through a list of names. The names are in a circle and you can move clockwise or counter-clockwise through the cirle.

```{r}
# read in
dat_1.2 <- read.table(here::here("everybody_codes_e2025_q01_p2.txt"),
                    sep = "")


substr(names_1$names, 1, 1)
```

```{r}
```

```{r}
#| echo: false
2 * 2
```
